---
title: "SLOO_BSB"
author: "Margaret Walker"
date: "March 30, 2016"
output: pdf_document
---
```{r}
library(geoR)
library(fields)
```

Bringing in BSB data 
```{r}
data <- read.csv("./Data/SpatialBSB.csv")
bsb <- subset(data, select=2:23)
```
Best model based on BIC from model selection 
```{r}
bsbbic <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
summary(bsbbic)
```
Determining split distance 
Examining spatial dependnece at the trap level by plotting the distance of the residuals vs the distance of all the points. 
```{r}
indices = sample(nrow(dat_sub), nrow(dat_sub) / 5)
dat_sub = data.frame(Lat =bsb$Lat, Long=bsb$Long, 
                     resid = residuals(bsbbic, 'response'))
dat_sub = dat_sub[indices, ]

xydist <- dist(dat_sub[ , c('Long', 'Lat')])
d_resid = dist(dat_sub$resid)

plot(d_resid ~ xydist, type='n')
lines(lowess(xydist, d_resid), lwd=2, col='red')

```
appears there is no real spatial autocorrelation at the trap scale.  Will go forward with tradiational cross validation. See CrossValidation.rmd  
