---
title: "SpatialDataExploration"
author: "Margaret Walker"
date: "March 28, 2016"
output: pdf_document
---
#Data exploration and manipuation for thesis data
This markdown contains data manipulation from raw chevron trap data to data to be used in model selection for black sea bass and white grunt. 

1.
Filtering raw chevron trap data into smaller datasets. 
Should already be subsetted for valid catch codes and normal sample duration. 
First, selecting just the columns I will need and saving as csv.
```{r}
Raw.Chevron.Trap.Data <- read.csv("./Data/Raw Chevron Trap Data.csv")
rawchevron <- data.frame(Raw.Chevron.Trap.Data)
allraw <- subset(rawchevron, select = c(Event, PCG, Year, Lat, Long, Depth, Duration, Temp, Salinity, Black.Sea.Bass.Abund, Red.Porgy.Abund, Red.Snapper.Abund, Scamp.Abund, Stenotomus.sp.Abund, Tomtate.Abund, White.Grunt.Abund ))
library(xlsx)
write.csv(allraw, "RawChevron.csv")
```
Subsetting for black sea bass. 
Add presence absence column for black sea bass.
Make a csv for just 2011-2015.
```{r}
allbsb <- subset(allraw, select =c(Event, PCG, Year, Lat, Long, Depth, Duration, Temp, Salinity, Black.Sea.Bass.Abund, Red.Snapper.Abund, Scamp.Abund, Stenotomus.sp.Abund, Tomtate.Abund))
allbsb$Presence_absence <- ifelse(allbsb$Black.Sea.Bass.Abund!=0, 1,0)
write.csv(allbsb, "rawBSB.csv")
newbsb <- subset(allbsb, Year >= 2011) 
newbsb <- na.exclude(newbsb)
write.csv(newbsb, "newBSB.csv")
```
Subsetting for white grunt. 
Add presence absence column for white grunt
Make a csv for just 2011-2015. 
```{r}
allwg <- subset(allraw, select =c(Event, PCG, Year, Lat, Long, Depth, Duration, Temp, Salinity, Red.Porgy.Abund, Red.Snapper.Abund, White.Grunt.Abund))
allwg$Presence_absence <- ifelse(allwg$White.Grunt.Abund!=0, 1, 0)
write.csv(allwg, "rawWG.csv")
newwg <- subset(allwg, Year >= 2011)
write.csv(newwg, "newWG.csv")
newwg <- na.exclude(newwg)
```
Using Joey's spatial smoothing code to get habitat data for all of my samples. 
First for BSB.
```{r}
set.seed(1234)
library(RColorBrewer)
library(maptools)
library(gridBase)
library(grid)
library(rgdal)
library(fields)
```
First, bring in video data and use weighted k nearest neighbor approach to extrapolate habitat value based on lat, long for my data.  
```{r}
Video.Data <- read.csv("./Data/Video Data.csv")
Video.Data <- Video.Data[,-1]

Habitat.Data <- subset(Video.Data,Valid_Video=="Yes")
#Habitat.Data <- subset(Habitat.Data,Station_Code!="Recon")
Habitat.Data <- Habitat.Data[,c(2,12,13,72:78)]

library(plyr)
Habitat.Data$Biota_Height <- ordered(Habitat.Data$Biota_Height,
                                     levels=c("Unknown",
                                              "No attached biota",
                                              "Max height < 0.5 m",
                                              "Max Height >0.5 m"))
Habitat.Data$Biota_Height <- with(Habitat.Data,revalue(Biota_Height,c("Unknown"=NA)))
Habitat.Data$Substrate_Size <- ordered(Habitat.Data$Substrate_Size,
                                       levels=c("Substrate cannot be seen",
                                                "No consolidated sediment",
                                                "> 50% consolidated sediment <1 m in diameter",
                                                ">50% Consolidated Sediment > 1 m in diameter"))
Habitat.Data$Substrate_Size <- with(Habitat.Data,revalue(Substrate_Size,
                                                         c("Substrate cannot 
                                                           be seen"=NA)))
Habitat.Data$Substrate_Relief <- ordered(Habitat.Data$Substrate_Relief,
                                         levels=c("Relief cannot be seen",
                                                  "<0.3 m of Relief",
                                                  "0.3 - 1 m of Relief",
                                                  ">1 m of Relief"))
Habitat.Data$Substrate_Relief <- with(Habitat.Data,revalue(Substrate_Relief,
                                                           c("Relief cannot
                                                             be seen"=NA)))
Habitat.Data <- droplevels(Habitat.Data)
BD.data <- na.exclude(Habitat.Data[,c("Latitude","Longitude", "Max_Biota_Density", "Mean_Biota_Density")])
SD.data <- na.exclude(Habitat.Data[,c("Latitude", "Longitude", "Max_Substrate_Density", "Mean_Substrate_Density")])
BH.data <- na.exclude(Habitat.Data[,c("Latitude", "Longitude", "Biota_Height")])
SS.data <- na.exclude(Habitat.Data[,c("Latitude", "Longitude", "Substrate_Size")])
SR.data <- na.exclude(Habitat.Data[,c("Latitude", "Longitude", "Substrate_Relief")])
```
Functions from Joey's script:
```{r}
# Function to test if the minimum distance between a point is > 1km from all points in another vector
dist.calc.function <- function(X,data,dist=1) {
    ifelse(min(rdist.earth(data.frame(Longitude=X[1],
                                      Latitude=X[2]),
                           data[,c("Longitude","Latitude")]))>dist,
           FALSE,
           TRUE)
}

# A useful function
  dircos<- function(x1){
             coslat1 <- cos((x1[, 2] * pi)/180)
             sinlat1 <- sin((x1[, 2] * pi)/180)
             coslon1 <- cos((x1[, 1] * pi)/180)
             sinlon1 <- sin((x1[, 1] * pi)/180)
             cbind(coslon1*coslat1, sinlon1*coslat1, sinlat1)}
colfunc <- colorRampPalette(c("royalblue","springgreen","red"))

kknn.fitted.values <- function(data) {
    subset(do.call("rbind",
                   lapply(data$fitted.values,
                          function(X) {
                              data.frame(value=X,
                                         type=attr(X,"kernel"),
                                         k=attr(X,"k"))
                              })),
           type==data$best.parameters$kernel&k==data$best.parameters$k)
}

```

Creating a grid that can be used for prediction using our spatial models later in the script
```{r}
xg <- make.surface.grid(fields.x.to.grid(Habitat.Data[,c("Longitude","Latitude")],
                                         nx=200,ny=200))
test <- sapply(lapply(1:nrow(xg), function(i) xg[i,]),dist.calc.function,data=Habitat.Data,dist=1.5) 
# Test whether individual grid points are closer than specified distance (in km) from a data point.
# Subsetting the grid to only include those grid values that are whithin a certain distance of an observed data point
xg <- data.frame(xg)
xg$test <- test
xg <- subset(xg,test==TRUE)
xg <- xg[,-3]
```

# Weighted k-Nearest Neighbors Models
## Maximum Biota Density
### No Covariates
first for BSB then WG. Creating maximum biota density column for white grunt and black sea bass data sets.
```{r}
library(kknn)
test <- data.frame(dircos(BD.data[,c("Longitude", "Latitude")]))
test$Max_Biota_Density <- BD.data$Max_Biota_Density
Max.BD.kknn.NoCov <- train.kknn(Max_Biota_Density~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); Max.BD.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$Max.BD.kknn <- predict(Max.BD.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1))  #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long", "Lat")]))
newwg$Max.BD.kknn <-predict(Max.BD.kknn.NoCov, newdata = list(sinlat1 =predict.df2$sinlat1))  #white grunt
```
Mean biota density
First for black sea bass then white grunt.
Creating mean biota desnity column for white grunt and black sea bass datasets. 
```{r}
test <- data.frame(dircos(BD.data[,c("Longitude", "Latitude")]))
test$Mean_Biota_Density <- BD.data$Mean_Biota_Density
Mean.BD.kknn.NoCov <- train.kknn(Mean_Biota_Density~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); Mean.BD.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$Mean.BD.kknn <- predict(Mean.BD.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1))  #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$Mean.BD.kknn <- predict(Mean.BD.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Biota Height
First for black sea bass then white grunt
Creating biota height column for black sea bass and white grunt datasets.
```{r}
test <- data.frame(dircos(BH.data[,c("Longitude", "Latitude")]))
test$Biota_Height <- BH.data$Biota_Height
BH.kknn.NoCov <- train.kknn(Biota_Height~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); BH.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$BH.kknn <- predict(BH.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1)) #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$BH.kknn <- predict(BH.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Maximum Substrate Density
Creating maximum substrate density column for black sea bass and white grunt data sets. 
```{r}
test <- data.frame(dircos(SD.data[,c("Longitude", "Latitude")]))
test$Max_Substrate_Density <- SD.data$Max_Substrate_Density
Max.SD.kknn.NoCov <- train.kknn(Max_Substrate_Density~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); Max.SD.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$Max.SD.kknn <- predict(Max.SD.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1)) #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$Max.SD.kknn <- predict(Max.SD.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Mean Substrate Density
Creating mean substrate density columns for black sea bass and white grunt datasets. 

```{r}
test <- data.frame(dircos(SD.data[,c("Longitude", "Latitude")]))
test$Mean_Substrate_Density <- SD.data$Mean_Substrate_Density
Mean.SD.kknn.NoCov <- train.kknn(Mean_Substrate_Density~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); Mean.SD.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$Mean.SD.kknn <- predict(Mean.SD.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1)) #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$Mean.SD.kknn <- predict(Mean.SD.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Substrate Size
Creating substrate size column for black sea bass and white grunt data sets. 
```{r}
test <- data.frame(dircos(SS.data[,c("Longitude", "Latitude")]))
test$Substrate_Size <- SS.data$Substrate_Size
SS.kknn.NoCov <- train.kknn(Substrate_Size~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); SS.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$SS.kknn <- predict(SS.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1)) #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$SS.kknn <- predict(SS.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Substrate Relief
Creating substrate relief column for black sea bass and white grunt data sets.
```{r}
test <- data.frame(dircos(SR.data[,c("Longitude", "Latitude")]))
test$Substrate_Relief <- SR.data$Substrate_Relief
SR.kknn.NoCov <- train.kknn(Substrate_Relief~sinlat1,
                         test,
                         kmax=40,
                         distance=2,
                         kernel=c("triangular",
                                  "epanechnikov",
                                  "biweight",
                                  "triweight",
                                  "optimal")); SR.kknn.NoCov
predict.df <- data.frame(dircos(newbsb[,c("Long","Lat")]))
newbsb$SR.kknn <- predict(SR.kknn.NoCov, newdata = list(sinlat1 = predict.df$sinlat1)) #black sea bass
predict.df2 <- data.frame(dircos(newwg[,c("Long","Lat")]))
newwg$SR.kknn <- predict(SR.kknn.NoCov, newdata = list(sinlat1 = predict.df2$sinlat1)) #white grunt
```
Creating new csv for bsb and wg with habiat variables. 

```{r}
write.csv(newbsb, "SpatialBSB.csv")
write.csv(newwg, "SpatialWG.csv")
```

