---
title: "SpatialModeling"
author: "Margaret Walker"
date: "March 29, 2016"
output: pdf_document
---
#Model Selection 

Bring in black sea bass data set and select just the necessary columns. 
```{r}
data <- read.csv("./Data/SpatialBSB.csv")
bsb <- subset(data, select=2:23)
```
Checking for correlation between variables: 
```{r}
plot(Mean.BD.kknn~Mean.SD.kknn, data=bsb) #correlated so only include biota density not both 
```
Also only include Mean.BD.kknn not Max.BD.kknn because that is the average value for each trap. 

Automated model selection process is below:
First we will run this on black sea bass for 2011-2015. 
The potential variables to be included in the model are as follows:
Year, Latitude, Depth, Duration, Temperature, Stenotomus sp. abundance, Red Snapper Abundance, Scamp abundance, Tomtate abundance, mean biota density, biota height, substrate size, and substrate relief. 
```{r}
library(MuMIn)
GLM.Names <- function(x,Names,familyabv) {
    paste0(familyabv,paste0(Names,x,collapse=""))
}
GLM.Formulas <- function(X,Y,offset=NULL,VarType,Names) {
    tmp <- X
    for(j in 1:length(tmp)) {
        if(VarType[j]=="F") {
            tmp[j] <- ifelse(tmp[j]==0,NA,Names[j])
            } else {
                if(VarType[j]=="C") {
                    tmp[j] <- ifelse(tmp[j]==0,NA,paste0("poly(",Names[j],",",tmp[j],",raw=TRUE)"))
                    } else {
                        stop("Invalid value for 'VarType' provided")
                    }
            }
        }
    if(is.null(offset)==FALSE) {
        offset <- paste0("+offset(",offset,")")
        } else {
            offset <- "+NA"
            }
    tmp2 <- paste0(Y,"~",offset,paste(tmp,collapse="+"))
    tmp3 <- gsub("+NA","",tmp2,fixed=TRUE)
    gsub("~NA+","~",tmp3,fixed=TRUE)
}
model.selection<-function(x,data){
  tmp<-try(glm(as.formula(x[2]),data=data,family=binomial))
  w<-ifelse(class(try(AIC(tmp)))=="numeric",AIC(tmp),NA)
  y<-ifelse(class(try(AICc(tmp)))=="numeric",AICc(tmp),NA)
  z<-ifelse(class(try(BIC(tmp)))=="numeric",BIC(tmp),NA)
  disp<-ifelse(class(try(AIC(tmp)))=="numeric",sum(resid(tmp,type="pearson")^2)/tmp$df.resid,NA)
  c(w,y,z,disp)
}
var.names<-c("Year","Lat","Depth","Duration", "Temp","Stenotomus.sp.Abund","Red.Snapper.Abund","Scamp.Abund","Tomtate.Abund", "Mean.BD.kknn", "BH.kknn", "SS.kknn", "SR.kknn")
variables<-c("Yr","Lat","Depth","Dur","Temp","Steno","RS","Scamp","Tom", "BD", "BH", "SS", "SR")
comb<-as.matrix(expand.grid(var1=c(0:1),var2=c(0:1),var3=c(0:1), var4=c(0:1),var5=c(0:1),var6=c(0:1),var7=c(0:1),var8=c(0:1),var9=c(0:1), var10=c(0:1),var11=c(0:1), var12=c(0:1), var13=c(0:1)))
count<-nrow(comb)
results<-list(name=c(),formula=c())
system.time(results$name<-apply(comb,1,GLM.Names,Names=variables,familyabv="bin"))
system.time(results$formula<-apply(comb,1,GLM.Formulas,Y="Presence_absence",Names=var.names,VarType=c("F","C","C","C","C","C","C","C","C", "C", "F", "F", "F")))
results.frame<-as.data.frame(results)
results.frame<-results.frame[-1,]
i<-sapply(results.frame,is.factor)
results.frame[i]<-lapply(results.frame[i],as.character)
results.frame$n<-seq(1,length(results.frame[,1]))
ptm<-proc.time()
bsb$Year<-as.factor(bsb$Year)
bin.ic<-t(apply(results.frame,1,FUN=model.selection,data=bsb))
rownames(bin.ic) <-results.frame[,1]
colnames(bin.ic) <- c("AIC","AICc","BIC","Dispersion")
bin.ic.Frame <- as.data.frame(bin.ic)
bin.ic.Frame$AIC.diff <- bin.ic.Frame$AIC-min(bin.ic.Frame$AIC,na.rm=TRUE)
bin.ic.Frame$AICc.diff <- bin.ic.Frame$AICc-min(bin.ic.Frame$AICc,na.rm=TRUE)
bin.ic.Frame$BIC.diff <- bin.ic.Frame$BIC-min(bin.ic.Frame$BIC,na.rm=TRUE)
bin.ic.BIC <- subset(bin.ic.Frame,BIC.diff<=10)
bin.ic.BIC <- bin.ic.BIC[with(bin.ic.BIC,order(BIC)),]; bin.ic.BIC
bin.ic.AIC <- subset(bin.ic.Frame,AIC.diff<=10)
bin.ic.AIC <- bin.ic.AIC[with(bin.ic.AIC,order(AIC)),]; bin.ic.AIC
bin.ic.AICc <- subset(bin.ic.Frame,AICc.diff<=5)
bin.ic.AICc <- bin.ic.AICc[with(bin.ic.AICc,order(AICc)),]; bin.ic.AICc
proc.time()-ptm
```
Best model according to BIC and AIC:
Plots created to look at the effects of the individual variables on probability of occurence using visreg. 
BIC model contains 9 variables while the AIC model contains 13 variables. 
```{r}
library(visreg)
bsbbic <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
visreg(bsbbic,type="conditional", scale="response", ylab="Probability of Occurrence")
bsbaic <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + Scamp.Abund + Mean.BD.kknn + BH.kknn + SS.kknn + SR.kknn, data=bsb, family="binomial" )
```
Removing individual variables to examine effects on BIC:
```{r}
minus_year <- glm(Presence_absence ~ Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_year)
minus_lat <- glm(Presence_absence ~ Year + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_lat)
minus_depth <- glm(Presence_absence ~ Year + Lat + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_depth)
minus_dur <- glm(Presence_absence ~ Year + Lat + Depth + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_dur)
minus_temp <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_temp)
minus_steno <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp +  Red.Snapper.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_steno)
minus_red <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Tomtate.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_red)
minus_tom <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + SR.kknn, data=bsb, family="binomial")
BIC(minus_tom)
minus_sr <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Stenotomus.sp.Abund + Red.Snapper.Abund + Tomtate.Abund, data=bsb, family="binomial")
BIC(minus_sr)
```



Now repeating the process for white grunt. 
First, brining in the data and selecting the necessary columns. 
```{r}
dat <- read.csv("./Data/SpatialWG.csv")
wg <- subset(dat, select=2:21)
```
Model selection:
Potential variables to be included in the model are as follows: 
Year, Latitude, Depth, Temperature, Duration, Red Porgy Abundance, red snapper abundance, mean biota density, substrate relief, substrate size, and biota height. 

```{r}
library(MuMIn)
GLM.Names <- function(x,Names,familyabv) {
    paste0(familyabv,paste0(Names,x,collapse=""))
}
GLM.Formulas <- function(X,Y,offset=NULL,VarType,Names) {
    tmp <- X
    for(j in 1:length(tmp)) {
        if(VarType[j]=="F") {
            tmp[j] <- ifelse(tmp[j]==0,NA,Names[j])
            } else {
                if(VarType[j]=="C") {
                    tmp[j] <- ifelse(tmp[j]==0,NA,paste0("poly(",Names[j],",",tmp[j],",raw=TRUE)"))
                    } else {
                        stop("Invalid value for 'VarType' provided")
                    }
            }
        }
    if(is.null(offset)==FALSE) {
        offset <- paste0("+offset(",offset,")")
        } else {
            offset <- "+NA"
            }
    tmp2 <- paste0(Y,"~",offset,paste(tmp,collapse="+"))
    tmp3 <- gsub("+NA","",tmp2,fixed=TRUE)
    gsub("~NA+","~",tmp3,fixed=TRUE)
}
model.selection<-function(x,data){
  tmp<-try(glm(as.formula(x[2]),data=data,family=binomial))
  w<-ifelse(class(try(AIC(tmp)))=="numeric",AIC(tmp),NA)
  y<-ifelse(class(try(AICc(tmp)))=="numeric",AICc(tmp),NA)
  z<-ifelse(class(try(BIC(tmp)))=="numeric",BIC(tmp),NA)
  disp<-ifelse(class(try(AIC(tmp)))=="numeric",sum(resid(tmp,type="pearson")^2)/tmp$df.resid,NA)
  c(w,y,z,disp)
}
var.names<-c("Year","Lat","Depth","Duration", "Temp","Red.Porgy.Abund","Red.Snapper.Abund", "Mean.BD.kknn", "BH.kknn", "SS.kknn", "SR.kknn")
variables<-c("Yr","Lat","Depth","Dur","Temp","RP","RS", "BD", "BH", "SS", "SR")
comb<-as.matrix(expand.grid(var1=c(0:1),var2=c(0:1),var3=c(0:1), var4=c(0:1),var5=c(0:1),var6=c(0:1),var7=c(0:1),var8=c(0:1),var9=c(0:1), var10=c(0:1),var11=c(0:1)))
count<-nrow(comb)
results<-list(name=c(),formula=c())
system.time(results$name<-apply(comb,1,GLM.Names,Names=variables,familyabv="bin"))
system.time(results$formula<-apply(comb,1,GLM.Formulas,Y="Presence_absence",Names=var.names,VarType=c("F","C","C","C","C","C","C","C","F", "F", "F")))
results.frame<-as.data.frame(results)
results.frame<-results.frame[-1,]
i<-sapply(results.frame,is.factor)
results.frame[i]<-lapply(results.frame[i],as.character)
results.frame$n<-seq(1,length(results.frame[,1]))
ptm<-proc.time()
wg$Year<-as.factor(wg$Year)
bin.ic<-t(apply(results.frame,1,FUN=model.selection,data=wg))
rownames(bin.ic) <-results.frame[,1]
colnames(bin.ic) <- c("AIC","AICc","BIC","Dispersion")
bin.ic.Frame <- as.data.frame(bin.ic)
bin.ic.Frame$AIC.diff <- bin.ic.Frame$AIC-min(bin.ic.Frame$AIC,na.rm=TRUE)
bin.ic.Frame$AICc.diff <- bin.ic.Frame$AICc-min(bin.ic.Frame$AICc,na.rm=TRUE)
bin.ic.Frame$BIC.diff <- bin.ic.Frame$BIC-min(bin.ic.Frame$BIC,na.rm=TRUE)
bin.ic.BIC <- subset(bin.ic.Frame,BIC.diff<=10)
bin.ic.BIC <- bin.ic.BIC[with(bin.ic.BIC,order(BIC)),]; bin.ic.BIC
bin.ic.AIC <- subset(bin.ic.Frame,AIC.diff<=10)
bin.ic.AIC <- bin.ic.AIC[with(bin.ic.AIC,order(AIC)),]; bin.ic.AIC
bin.ic.AICc <- subset(bin.ic.Frame,AICc.diff<=5)
bin.ic.AICc <- bin.ic.AICc[with(bin.ic.AICc,order(AICc)),]; bin.ic.AICc
proc.time()-ptm
```
Best aic and bic models.
The BIC model contains 6 variables while the AIC model contains 10 variables. 
Effects of individual variables on occurence are visualized with visreg. 
```{r}
wgbic <- glm(Presence_absence ~ Year + Lat + Depth + Temp + Red.Porgy.Abund + Mean.BD.kknn, data=wg, family="binomial")
par(mfcol=c(1,1))
visreg(wgbic,type="conditional", scale="response", ylab="Probability of Occurrence")
BIC(wgbic)
wgaic <- glm(Presence_absence ~ Year + Lat + Depth + Duration + Temp + Red.Porgy.Abund + Red.Snapper.Abund + Mean.BD.kknn + SS.kknn + SR.kknn, data=wg, family="binomial" )
```
Removing indiviaul variables and looking at effect on BIC.
```{r}
wg_minus_year <- glm(Presence_absence ~ Lat + Depth + Temp + Red.Porgy.Abund + Mean.BD.kknn, data=wg, family="binomial")
BIC(wg_minus_year)
wg_minus_lat <- glm(Presence_absence ~ Year + Depth + Temp + Red.Porgy.Abund + Mean.BD.kknn, data=wg, family="binomial")
BIC(wg_minus_lat)
wg_minus_depth <- glm(Presence_absence ~ Year + Lat +Temp + Red.Porgy.Abund + Mean.BD.kknn, data=wg, family="binomial")
BIC(wg_minus_depth)
wg_minus_temp <- glm(Presence_absence ~ Year + Lat + Depth + Red.Porgy.Abund + Mean.BD.kknn, data=wg, family="binomial")
BIC(wg_minus_temp)
wg_minus_rp <- glm(Presence_absence ~ Year + Lat + Depth + Temp +  Mean.BD.kknn, data=wg, family="binomial")
BIC(wg_minus_rp)
wg_minus_bd <- glm(Presence_absence ~ Year + Lat + Depth + Temp + Red.Porgy.Abund , data=wg, family="binomial")
BIC(wg_minus_bd)
```


